\TheTitle

### 2019-11-29 Paper1/results.tex
last comments from Martin
corrections and remarks Jannick

The computational domain is a circle with a diameter of $40$ meters. In the center, the NACA0012 airfoil with a length of one meter is located. The spatial mesh is composed of a coarsely discretized far field and a finely resolved region around the airfoil, since we are interested in the flow solution at the airfoil. Altogether, the mesh consists of 22361 triangular elements.

\begin{figure}[h!]
\centering
	\begin{subfigure}{0.3\linewidth}
		\centering
				\includegraphics[scale=0.2]{figs/Euler1DPlots5/sc100_ERho.png}
		\label{fig:referenceSolutionsub1}
	\end{subfigure}%	
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
				\includegraphics[scale=0.2]{figs/Euler1DPlots5/sc100_VarRho.png}
		\label{fig:referenceSolutionsub2}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/naca_mesh.png}
		\label{fig:referenceSolutionsMesh}
	\end{subfigure}
	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and mesh used in the computation of all presented methods.}
	\label{fig:referenceSolution}
\end{figure}

@@

	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
		\caption{}
		\label{fig:sub33}
	\end{subfigure}%
	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
		\caption{}
		\label{fig:sub34}
	\end{subfigure}
	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row). All intrusive methods converge to a residual $\varepsilon=5e-6$ whereas all non-intrusive methods converge to a residual $\varepsilon=1e-7$. All computations are performed with with 5 MPI threads.}
	\label{fig:L2ErrorSolution}
\end{figure}
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_ERho.png}
		\label{fig:sub1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/sg_4_9_ERho.png}
		\label{fig:sub2}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/osIPM4-4_ERho.png}
		\label{fig:sub1}
	\end{subfigure}\\
	\vspace{-0.35cm}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_VarRho.png}
		\label{fig:sub1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/sg_4_9_VarRho.png}
		\label{fig:sub2}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/osIPM4-4_VarRho.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the reference solution shown in Figure \ref{fig:referenceSolution}.}
	\label{fig:ERhoVarRho5}
\end{figure}

@@

Let us now compare results obtained with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order $9$. Note that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well. All adaptive methods use gPC polynomials of order 2 to 9 (i.e. 3 to 10 moments). Order $2$ uses $5$ quadrature points, orders $3$ to $6$ use $9$ quadrature points and orders $8$ and $9$ use $17$ quadrature points. When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the adaptive methods decrease the truncation order, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased. The remaining methods have been computed with $17$ quadrature points. The iteration in pseudo-time is performed until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.

First, let us mention that the adaptive SG method fails, since it yields negative densities during the iteration. The standard SG method however preserves positivity of mass, energy and pressure. The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation value and in Figure~\ref{fig:sub32} for the variance. When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:
\begin{itemize}
\item Compared to IPM, stochastic-Galerkin comes at a significantly reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
\item SG$_9$ shows a smaller error compared to IPM$_9$ for the expectation value. For the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
\end{itemize}
The proposed acceleration techniques show the following behavior:
\begin{itemize}
\item The One-Shot IPM (osIPM) method proposed in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the One-Shot idea, the method is denoted by \textit{adaptive One-Shot IPM} (adosIPM). This method reaches the steady state IPM solution with a faster runtime than SG.
\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by retardation adosIPM (readosIPM), which further decreases runtime. Here, we use two different strategies to increase the accuracy: First, we steadily increase the maximal truncation order when the residual approaches zero. To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order. The residual values are then determined to be $6\cdot 10^{-5},3\cdot 10^{-5},2.2\cdot 10^{-5}$ and $2\cdot 10^{-5}$. The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$. Strategy 1 is depicted in purple, Strategy 2 is depicted in yellow. It can be seen that both approaches reach the IPM$_9$ error for the same run time. Hence, we deduce that a naive choice of the refinement retardation strategy suffices to yield a satisfactory behavior.
\end{itemize}

Let us now compare the results from intrusive methods with those of SC. For every quadrature point, SC iterates the solution until the density lies below a threshold of $\varepsilon = 1e^{-7}$. Recording the error of intrusive methods during the iteration is straight forward. To record the error of SC, we couple all quadrature points after each iteration to evaluate the expectation value and variance, which destroys the black-box nature and results in additional costs. The collocation runtimes that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D} are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error. Our stochastic-Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e. $5,9$ and $17$ quadrature points. The comparison of intrusive methods with non-intrusive methods shows the following:
\begin{itemize}
\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1e^{-7}$ compared to $\varepsilon = 5e^{-6}$ for intrusive methods).
\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns. Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
\end{itemize}

@@

In the following, we assume a fixed angle of attack with $\phi = 1.25$ degrees and study the effect of two sources of uncertainties, namely the farfield pressure and Mach number. The farfield pressure is ${p \sim U(100\;325,102\;325)}$ Pa and the Mach number is $Ma \sim U(0.775,0.825)$. Since this problem has only a two-dimensional uncertainty, we use a tensorized quadrature set, which in our experiments proved to be more efficient than a sparse grid quadrature. For SC, we use quadrature sets with $5^2$, $9^2$ and $17^2$ quadrature nodes and compare against SG with moments up to total degree $9$ as well as adaptive IPM with refinement retardation (with and without the One-Shot strategy). The IPM method uses moment orders ranging from $1$ to $9$ adaptively with refinement barriers of $\delta_{-} = 1\cdot 10^{-4}$ and $\delta_{+} = 1\cdot 10^{-5}$. The refinement retardation allows the truncation order to have a maximal total degree of $1$ until a residual of $\varepsilon = 1.5\cdot 10^{-5}$ and then increases the truncation order by one when the residual is reduced by an amount of $5\cdot 10^{-6}$. Hence, the maximal truncation order of $9$ is reached when the residual is below $\varepsilon = 7\cdot 10^{-6}$. Again, the refinement strategy let to negative densities for SG resulting in a failure of the method.
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_E[rho].pdf}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_Var[rho].pdf}
		\label{fig:sub2}
	\end{subfigure}

@@

The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}. To determine the error, we used Stochastic-Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule. As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method. Furthermore, the error obtained with the intrusive methods requires a total degree of $M = 9$, i.e. $N = 55$ moments to reach the error level of SC with 17 quadrature points per dimension i.e. $17^2 = 289$ collocation points. Note that this effect has also been observed in the one-dimensional case, however now for multi-dimensional problems, the reduced number of unknowns needed for intrusive methods to obtain a certain error weighs in more heavily. This shows one promising characteristic of intrusive techniques and points to their applicability for higher dimensional problems. In contrast to before, the SG error level is smaller than the IPM error for both, the expectation value and variance. Furthermore, when comparing IPM with and without One-Shot, one can observe that the effect of this acceleration strategy weighs in more heavily than it did for the one-dimensional case. This behavior is expected, since every iterate of the optimization problem becomes more expensive when the dimension is increased. This means, using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the One-Shot strategy to be an effective approach, especially for problems with high uncertain dimension.

@@

When looking at the computed IPM expectation value and variance (see Figure~\ref{fig:reados2DE} and \ref{fig:reados2DVar}) and comparing the results against the reference solution in Figure~\ref{fig:referenceSolutionsub2DE} and \ref{fig:referenceSolutionsub2DVar}, one can observe that since the uncertainty effects the result more heavily than in the one-dimensional case, the IPM solution still shows a step-like profile. However, we are able to capture the main features of the reference solution. The increased effect of the uncertainty can again be seen in Figure~\ref{fig:readosIPMEVar2DRefinement}, where a refinement level of $8$, i.e. a truncation order of $9$ is chosen on a larger portion than for the one-dimensional case.

>>>>

The computational domain is a circle with a diameter of $40$ meters. In the center, the NACA0012 airfoil with a length of one meter is located. The spatial mesh is composed of a coarsely discretized far field and a finely resolved region around the airfoil, since we are interested in the flow solution at the airfoil. Altogether, the mesh consists of 22361 triangular elements. \remja{Note, that the cell density of the mesh was adjusted in order to capture areas of high variance and therefore the mesh is significantly different to a mesh classically applied in computational fluid dynamics.}

\begin{figure}[h!]
\centering
	\begin{subfigure}{0.3\linewidth}
		\centering
				\includegraphics[scale=0.2]{figs/Euler1DPlots5/sc100_ERho.png}
		\label{fig:referenceSolutionsub1}
	\end{subfigure}%	
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
				\includegraphics[scale=0.2]{figs/Euler1DPlots5/sc100_VarRho.png}
		\label{fig:referenceSolutionsub2}
	\end{subfigure}
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/naca_mesh.png}
		\label{fig:referenceSolutionsMesh}
	\end{subfigure}
	\caption{Reference solution E$[\rho]$ and Var$[\rho]$ and the mesh close to the airfoil which is used in the computation of all presented methods.}
	\label{fig:referenceSolution}
\end{figure}

@@

	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_E[rho]_sc.pdf}
		\caption{}
		\label{fig:sub33}
	\end{subfigure}%
	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[scale=0.55]{figs/errorEuler/L2_error_Var[rho]_sc.pdf}
		\caption{}
		\label{fig:sub34}
	\end{subfigure}
	\caption{Comparison of the relative L$^2$-error for the density for IPM related methods (first row) and of the best performing IPM method in comparison with SC (second row). All intrusive methods converge to a residual $\varepsilon=5\cdot 10^{-6}$, whereas all non-intrusive methods converge to a residual $\varepsilon=\cdot 10^{-7}$. All computations are performed with with 5 MPI threads.}
	\label{fig:L2ErrorSolution}
\end{figure}
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_ERho.png}
		\label{fig:sub1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/sg_4_9_ERho.png}
		\label{fig:sub2}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/osIPM4-4_ERho.png}
		\label{fig:sub1}
	\end{subfigure}\\
	\vspace{-0.35cm}
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/EulerSC5-2Res1e-6_VarRho.png}
		\label{fig:sub1}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/sg_4_9_VarRho.png}
		\label{fig:sub2}
	\end{subfigure}%
	\hfill
	\begin{subfigure}{0.3\linewidth}
		\centering
		\includegraphics[scale=0.2]{figs/Euler1DPlots5/osIPM4-4_VarRho.png}
		\label{fig:sub1}
	\end{subfigure}
	\caption{E$[\rho]$ and Var$[\rho]$ computed with SC$^5$, SG$_4$, IPM$_4$ (from left to right). Compare to the reference solution shown in Figure \ref{fig:referenceSolution}.}
	\label{fig:ERhoVarRho5}
\end{figure}

@@

Let us now compare results obtained with stochastic-Galerkin and IPM as well as its proposed acceleration techniques at a fixed moment order $9$. Note, that since IPM generalizes SG, all proposed techniques can be used for stochastic-Galerkin as well. All adaptive methods use gPC polynomials of order 2 to 9 (i.e. 3 to 10 moments). Order $2$ uses $5$ quadrature points, orders $3$ to $6$ use $9$ quadrature points and orders $8$ and $9$ use $17$ quadrature points. When the smoothness indicator \eqref{eq:errorIndicator} lies below $\delta_{-} = 2\cdot 10^{-4}$, the adaptive methods decrease the truncation order, if it lies above $\delta_{+} = 2\cdot 10^{-5}$ the truncation order is increased. The remaining methods have been computed with $17$ quadrature points. The iteration in pseudo-time is performed until the expectation value of the density fulfills the stopping criterion \eqref{eq:residualSteady} with $\varepsilon = 5\cdot 10^{-6}$, however it can be seen that the error saturates already at a bigger residual.

First, let us mention that the adaptive SG method fails, since it yields negative densities during the iteration. The standard SG method however preserves positivity of mass, energy and pressure. The change of the relative L$^2$-error during the iteration to the steady state has been recorded in Figure~\ref{fig:sub31} for the expectation value and in Figure~\ref{fig:sub32} for the variance. When comparing intrusive methods without acceleration techniques as well as SC, the following properties emerge:
\begin{itemize}
\item Compared to IPM, stochastic-Galerkin comes at a significantly reduced runtime, meaning that the IPM optimization problem requires a significant computational effort.
\item For the expectation value SG$_9$ shows a smaller error compared to IPM$_9$, while for the variance, we see the opposite, i.e. IPM yields a better solution approximation than SG.
\end{itemize}
The proposed acceleration techniques show the following behavior:
\begin{itemize}
\item The One-Shot IPM (osIPM) method proposed in Section~\ref{sec:OneShotIPM} reduces the runtime while yielding the same error as the classical IPM method.
\item When using adaptivity (see Algorithm~\ref{alg:ad-IPM}) in combination with the One-Shot idea, the method is denoted by \textit{adaptive One-Shot IPM} (adosIPM). This method reaches the steady state IPM solution with a faster runtime than SG.
\item The idea of refinement retardation combined with adosIPM (see Algorithm~\ref{alg:adosIPM}) is denoted by \textit{retardation adosIPM} (readosIPM), which further decreases runtime. Here, we use two different strategies to increase the accuracy: First, we steadily increase the maximal truncation order when the residual approaches zero. To determine residual values for a given set of truncation orders $2,4,5$ and $8$, we study at which residual level the IPM method reaches a saturated error level for each truncation order. The residual values are then determined to be $6\cdot 10^{-5},3\cdot 10^{-5},2.2\cdot 10^{-5}$ and $2\cdot 10^{-5}$. The second, straight forward strategy converges the solution on a low truncation order of $2$ to a residual of $10^{-5}$ and then switches to a maximal truncation order of $9$. Strategy 1 is depicted in purple, Strategy 2 is depicted in yellow. It can be seen that both approaches reach the IPM$_9$ error for the same run time. Hence, we deduce that a naive choice of the refinement retardation strategy suffices to yield a satisfactory behavior.
\end{itemize}

Let us now compare the results from intrusive methods with those of SC. For every quadrature point, SC iterates the solution until the density lies below a threshold of $\varepsilon = 1\cdot 10^{-7}$. Recording the error of intrusive methods during the iteration is straight forward. To record the error of SC, we couple all quadrature points after each iteration to evaluate the expectation value and variance, which destroys the black-box nature and results in additional costs. The collocation runtimes, that are depicted on the $x$-axis in Figures~\ref{fig:ErrorDifferentQuad}, \ref{fig:sub33}, \ref{fig:sub34} and \ref{fig:L2ErrorSolution2D}, are however rescaled to the runtimes that we achieve when running the collocation code in its original, black-box framework without recording the error. Our stochastic-Collocation computations use Clenshaw-Curtis quadrature levels $2,3$ and $4$, i.e. $5,9$ and $17$ quadrature points. The comparison of intrusive methods with non-intrusive methods shows the following:
\begin{itemize}
\item SC requires a smaller residual to converge to a steady state solution (we converge the results to $\varepsilon = 1\cdot 10^{-7}$ compared to $\varepsilon = 5\cdot 10^{-6}$ for intrusive methods).
\item Again, intrusive methods yield improved solutions compared to SC with the same number of unknowns. Actually, the error obtained with 17 unknowns when using SC is comparable with the error obtained with 10 unknowns when using intrusive methods.
\end{itemize}

@@

In the following, we assume a fixed angle of attack with $\phi = 1.25$ degrees and study the effect of two sources of uncertainties, namely the farfield pressure and Mach number. The farfield pressure is ${p \sim U(100\;325,102\;325)}$ Pa and the Mach number is $Ma \sim U(0.775,0.825)$. Since this problem only has a two-dimensional uncertainty, we use a tensorized quadrature set, which in our experiments proved to be more efficient than a sparse grid quadrature. For SC, we use quadrature sets with $5^2$, $9^2$ and $17^2$ quadrature nodes and compare against SG with moments up to total degree $9$ as well as adaptive IPM with refinement retardation (with and without the One-Shot strategy). The IPM method uses moment orders ranging from $1$ to $9$ adaptively with refinement barriers $\delta_{-} = 1\cdot 10^{-4}$ and $\delta_{+} = 1\cdot 10^{-5}$. The refinement retardation allows the truncation order to have a maximal total degree of $1$ until a residual of $\varepsilon = 1.5\cdot 10^{-5}$ and then increases the truncation order by one when the residual is reduced by an amount of $5\cdot 10^{-6}$. Hence, the maximal truncation order of $9$ is reached when the residual is below $\varepsilon = 7\cdot 10^{-6}$. Again, the refinement strategy let to negative densities for SG resulting in a failure of the method.
\begin{figure}[h!]
\centering
	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_E[rho].pdf}
		\label{fig:sub1}
	\end{subfigure}%
	\begin{subfigure}{0.5\linewidth}
		\centering
				\includegraphics[scale=0.5]{figs/Euler2D/L2_error_Var[rho].pdf}
		\label{fig:sub2}
	\end{subfigure}

@@

The error during the iteration has been recorded in Figure~\ref{fig:L2ErrorSolution2D}. To determine the error, we used stochastic-Collocation with a $50$ by $50$ Gauss-Legendre quadrature rule. As in the one-dimensional NACA testcase, the acceleration techniques lead to a heavily reduced runtime of the IPM method. Furthermore, the error obtained with the intrusive methods requires a total degree of $M = 9$, i.e. $N = 55$ moments to reach the error level of SC with 17 quadrature points per dimension i.e. $17^2 = 289$ collocation points. Note that this effect has also been observed in the one-dimensional case, however now for multi-dimensional problems, the reduced number of unknowns required for intrusive methods to obtain a certain error becomes more apparent. This shows one promising characteristic of intrusive techniques and points to their applicability for higher dimensional problems. In contrast to before, the SG error level is smaller than the IPM error for both, the expectation value and variance. Furthermore, when comparing IPM with and without One-Shot, one can observe that the effect of this acceleration strategy weighs in more heavily than it did for the one-dimensional case. This behavior is expected, since every iterate of the optimization problem becomes more expensive when the dimension is increased. This means, using a method with less iterations for every computation of the dual variables heavily reduces computational costs, which is why we consider the One-Shot strategy to be an effective approach, especially for problems with high uncertain dimension.

@@

When looking at the computed IPM expectation value and variance (see Figure~\ref{fig:reados2DE} and \ref{fig:reados2DVar}) and comparing the results against the reference solution in Figure~\ref{fig:referenceSolutionsub2DE} and \ref{fig:referenceSolutionsub2DVar}, one can observe that since the uncertainties have a bigger effect on the result than in the one-dimensional case, the IPM solution still shows a step-like profile. However, we are able to capture the main features of the reference solution. The increased effect of the uncertainty can again be seen in Figure~\ref{fig:readosIPMEVar2DRefinement}, where a refinement level of $8$, i.e. a truncation order of $9$ is chosen on a larger portion than for the one-dimensional case. \remja{Visualizing the variance on a logarithmic scale in Figure~\ref{fig:referenceSolutionsub2DVarLog} further shows that the applied refinement indicators work well and use higher order moments in areas of high variance.}

### 2019-11-29 Paper1/summary_outlook.tex
corrections and remarks Jannick

\section{Summary and outlook}
\label{sec:summary_outlook}
In this work, we proposed acceleration techniques, which can be applied to the intrusive framework of IPM and SG: We use a One-Shot technique to iterate moments and their corresponding dual variables to their steady state simultaneously. By not fully converging the dual iteration, we can reduce computational costs. Additionally, we make use of adaptivity and propose to keep a low maximal truncation order for most of the iteration to the steady state solution. Since complicated structures in the uncertain dimension only appear on a small portion of the spacial mesh, we are able to heavily reduce computational costs. The effects of the proposed techniques have been demonstrated by comparing results obtained with IPM as well as SG against SC. In our test cases, the intrusive methods yield the same error level as SC for a reduced runtime, especially since intrusive methods require less unknowns to achieve a certain accuracy due to aliasing errors. In higher-dimensional problems, this effect is amplified since the number of unknowns to achieve a certain total degree is asymptotically smaller than the number of quadrature points in a tensorized or sparse grid (\eqref{eq:numberBasisFcts} instead of $O(M(\log_2(M)^{p-1}))$). Furthermore, we could observe that the required residual at which the solution of intrusive methods reaches a steady state is smaller than for SC. Additionally, the ability to adaptively change the truncation order helps intrusive methods to compete with SC in terms of computational runtime.

In future work, we aim at further accelerating the IPM method by using non-exact Hessian approximations. Similar to the One-Shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to a steady state. Hessian approximations that can be interesting are BFGS and sparse BFGS \cite[Chapter~6.1]{nocedal2006numerical}, which construct the Hessian from previously computed gradients. Note that this strategy will increase the used memory, since old Hessians or gradients from a certain number of old time steps need to be saved in every spacial cell.
Even though the non-intrusive nature of stochastic-Collocation or Monte Carlo methods allows an easy implementation, it can be important to intrusively modify the code in order to fully exploit all acceleration potentials. Synchronizing the time updates of the solution at different quadrature points yields an increased control over the solution during the computation, which can for example be uses to employ adaptive methods. In this case one can switch to a fine quadrature level in a certain spatial cell by for example computing moments with the given coarse set of collocation points. From these moments one can compute an IPM reconstruction, which one can then evaluate at a finer quadrature set. Another example of breaking up the non-intrusive nature of Monte Carlo methods can be found in \cite{poette2019gpc}, where the generation of random samples is combined with the sampling after collisions to increase efficiency. Furthermore, we aim at applying the proposed acceleration techniques to SG methods with hyperbolicity limiters \cite{wu2017stochastic,schlachter2018hyperbolicity}.

>>>>

\section{Summary and outlook}
\label{sec:summary_outlook}
In this work, we proposed acceleration techniques, which are applicable to the intrusive nature of IPM and SG: We use a One-Shot technique to iterate moments and their corresponding dual variables to their steady state simultaneously. By not fully converging the dual iteration, we can reduce computational costs. Additionally, we make use of adaptivity and propose to keep a low maximal truncation order for most of the iteration to the steady state solution. Since complicated structures in the uncertain dimension only appear on a small portion of the spacial mesh, we are able to heavily reduce computational costs. The effects of the proposed techniques have been demonstrated by comparing results obtained with IPM as well as SG against SC. In our test cases, the intrusive methods yield the same error level as SC for a reduced runtime, especially since intrusive methods require less unknowns to achieve a certain accuracy due to aliasing errors. In higher-dimensional problems, this effect is amplified since the number of unknowns to achieve a certain total degree is asymptotically smaller than the number of quadrature points in a tensorized or sparse grid (\eqref{eq:numberBasisFcts} instead of $O(M(\log_2(M)^{p-1}))$). Furthermore, we could observe that the required residual at which the solution of intrusive methods reaches a steady state is smaller than for SC. Additionally, the ability to adaptively change the truncation order helps intrusive methods to compete with SC in terms of computational runtime.

In future work, we aim at further accelerating the IPM method by using non-exact Hessian approximations. Similar to the One-Shot idea of not fully converging the dual problem, it seems to be plausible to not spend too much time on computing the Hessian when the moments are not close to a steady state. Hessian approximations that can be interesting are BFGS and sparse BFGS \cite[Chapter~6.1]{nocedal2006numerical}, which construct the Hessian from previously computed gradients. Note that this strategy will increase the used memory, since old Hessians or gradients from a certain number of old time steps need to be saved in every spacial cell.
Even though the non-intrusive nature of stochastic-Collocation or Monte Carlo methods allows an easy implementation, as shown in this work, it can be helpful to intrusively modify the code in order to fully exploit all acceleration potentials. Synchronizing the time updates of the solution at different quadrature points yields an increased control over the solution during the computation, which can for example be uses to employ adaptive methods. In this case one can switch to a fine quadrature level in a certain spatial cell by for example computing moments with the given coarse set of collocation points. From these moments one can compute an IPM reconstruction, which one can then evaluate at a finer quadrature set. Another example of breaking up the non-intrusive nature of Monte Carlo methods can be found in \cite{poette2019gpc}, where the generation of random samples is combined with the sampling after collisions to increase efficiency. Furthermore, we aim at applying the proposed acceleration techniques to SG methods with hyperbolicity limiters \cite{wu2017stochastic,schlachter2018hyperbolicity}.
